---
title: "Apartment Water Consumption Prediction Model"
output:
  html_document: default
  html_notebook: default
---

#Frame

Our apartment Complex has implemented individual water meters for all apartments since last two years. This has led to a lot of savings in cumulative water bill of the apartment complex since now each apartment resident can track his consumption clearly.  This is a marked difference from the earlier mechanism, where in the absence of indivdual readings the total bill supplied by the vendor was divided equally among all the apartments.

One of the common questions asked by interested tenants is the approx water bill they would incur in case they rent an apartment in our complex.

Can we predict approx water consumption for an interested party given some parameters about them?

# Acquire

Data for a typical month was provided by the manager of our apartment complex.  August last year was taken as a typical month (a lot of other months had vacations and other variations).  This data was formatted using excel (removing names of people) and saved as a csv file. Apartment complex has 72 apartments.


```{r Acquire}
#getwd()
water_consumption=read.csv("Apartment_water_consumption.csv",stringsAsFactors = FALSE)
str(water_consumption)

```
# Refine and Transform

1. Remove unwanted columns (We don't need columns like Sl no and Apartment No)
```{r }
water_consumption_cleaned=water_consumption[3:9]
str(water_consumption_cleaned)
```
2. Rename columns
```{r }
colnames(water_consumption_cleaned)=c("Resident_status","Family_Status",
                                      "Number_Adults","Number_Children","Number_Toilet","Area_of_apt","Water_Consumption") 
        
```
3. Change the data types (Change a few columns to factors)

```{r }
water_consumption_cleaned$Resident_status <- as.factor(water_consumption_cleaned$Resident_status)
water_consumption_cleaned$Family_Status <- as.factor(water_consumption_cleaned$Family_Status)
water_consumption_cleaned$Number_Toilet <- as.factor(water_consumption_cleaned$Number_Toilet)
str(water_consumption_cleaned)
```
4. Remove NA and Blanks. This will also remove the last three rows which are not apartments, but clubhouse, common area etc for which there are separate water meters.
```{r}

table(water_consumption_cleaned$Family_Status, useNA = "ifany")
water_consumption_cleaned=water_consumption_cleaned[complete.cases(water_consumption_cleaned),]
table(water_consumption_cleaned$Family_Status, useNA = "ifany")
str(water_consumption_cleaned)

```
#Explore
```{r}
library(ggplot2)
boxplot(water_consumption_cleaned$Water_Consumption~water_consumption_cleaned$Area_of_apt)
# We can clearly make out an outlier for an apartment for area 1255

qplot(Number_Adults, Water_Consumption, data=water_consumption_cleaned, shape=Number_Toilet, color=Number_Toilet, 
   facets=Family_Status~Resident_status, size=I(3),
   xlab="Number_Adults", ylab="Water_Consumption") 
```

#Model

```{r warning=FALSE}
# We will use five different models and do a comparison to get the best model
library(caret)
set.seed(5000)
# Multiple linear regression. We are taking the log of Water_Consumption as the relation between that and several variables are non -linear
m1 <- train(log(Water_Consumption) ~ ., data = water_consumption_cleaned, method = "lm")
# Regression Tree
m2 <-train(Water_Consumption ~ ., data = water_consumption_cleaned, method = "rpart", metric="Rsquared")
# Model Tree
m3 <-train(Water_Consumption ~ ., data = water_consumption_cleaned, method = "M5",metric="Rsquared")
# Random Forest
m4 <-train(Water_Consumption ~ ., data = water_consumption_cleaned, method = "rf",metric="Rsquared")

# Compare the models based on maximum R-Squared
RSquared_Linear_Regression=m1$results[which.max(m1$results$Rsquared),]["Rsquared"]
RSquared_Reg_Tree=m2$results[which.max(m2$results$Rsquared),]["Rsquared"]
RSquared_Model_Tree=m3$results[which.max(m3$results$Rsquared),]["Rsquared"]
RSquared_Random_Forest=m4$results[which.max(m4$results$Rsquared),]["Rsquared"]

library(data.table)

DT = data.table(ID = c("Linear_Reg","Reg_Tree","Model_Tree","Random_Forest"), R_Squared=c(RSquared_Linear_Regression,RSquared_Reg_Tree,RSquared_Model_Tree,RSquared_Random_Forest))

print(as.matrix(DT))

#From the R-Squared Table for the 4 models we can see that Linear_regression has the highest R-Squared. #However it is still low at approx 37%. Can we improve it. recall that from the visualiazation we did #during explore phase we found a big outlier. Can that be causing an issue? Let us remove that and try

water_consumption_cleaned_removed=water_consumption_cleaned[-which.max(water_consumption_cleaned$Water_Consumption),]

set.seed(5000)
# Multiple linear regression. We are taking the log of Water_Consumption as the relation between that and several variables are non -linear
m11 <- train(log(Water_Consumption) ~ ., data = water_consumption_cleaned_removed, method = "lm")
# Regression Tree
m21 <-train(Water_Consumption ~ ., data = water_consumption_cleaned_removed, method = "rpart", metric="Rsquared")
# Model Tree
m31 <-train(Water_Consumption ~ ., data = water_consumption_cleaned_removed, method = "M5",metric="Rsquared")
# Random Forest
m41 <-train(Water_Consumption ~ ., data = water_consumption_cleaned_removed, method = "rf",metric="Rsquared")

# Compare the models based on maximum R-Squared
RSquared_Linear_Regression_2=m11$results[which.max(m11$results$Rsquared),]["Rsquared"]
RSquared_Reg_Tree_2=m21$results[which.max(m21$results$Rsquared),]["Rsquared"]
RSquared_Model_Tree_2=m31$results[which.max(m31$results$Rsquared),]["Rsquared"]
RSquared_Random_Forest_2=m41$results[which.max(m41$results$Rsquared),]["Rsquared"]

DT2 = data.table(ID = c("Linear_Reg","Reg_Tree","Model_Tree","Random_Forest"), R_Squared=c(RSquared_Linear_Regression_2,RSquared_Reg_Tree_2,RSquared_Model_Tree_2,RSquared_Random_Forest_2))

print(as.matrix(DT2))

# There is some improvement with R-Squared of regression now above 41%

```
#Communicate
What could be the approximate consumption of a new interested tenant with the follwing characteristics.
Family of three (2 adults, 1 kid) , trying to rent a two bedroom apartment of area 1210 sqft



```{r, warning=FALSE}
newdata = data.frame(Area_of_apt=as.integer(1210),Number_Toilet=factor(2),Number_Adults=as.integer(2),
                     Number_Children=as.integer(1),
                     Family_Status="Yes",Resident_status="Tenant")
Predicted_Water_consumption=exp(predict(m11,newdata,interval="predict"))

#Predicted Monthly Consumption in liters

print(Predicted_Water_consumption)

```
#Insights

The R-Squared value is not very high.  And we need to test this on lot of test data to see the effectiveness of the prediction.  Additional predictors like WashingMachine, Reverse_Osmosis_Filter(which wastes lot of water) could be added. Still our model can give a useful indicator as to the approximate usage of monthly water.